\textsc{PAPERCLIP}: Associating Astronomical Observations and Natural Language with Multi-Modal Models

### 2024-03-07 paper/main.tex
Address jdt comments

\def\preprintno{XXXX} % Insert correct preprint number

@@

There is considerable interest in developing custom foundation models for the sciences \citep[e.g., ][]{batatia2023foundation,subramanian2023towards}, with astrophysics being ripe for such an effort given the availability of large amounts of publicly-available data and diverse modes of interacting with it.
%
The multi-modality inherent to astrophysical observations, with different types of data (e.g., images, spectra, light curves, textual descriptions) often available for a given target object, presents a unique opportunity.
%
This multi-modality was recently exploited in \textsc{AstroCLIP}~\citep{lanusse2023astroclip} to construct a joint embedding space between multi-band images and optical spectra from the Dark Energy Spectroscopic Instrument (DESI).
%
\textsc{AstroLLaMA}~\citep{nguyen2023astrollama,perkowski2024astrollama} is another recent effort to fine-tune a publicly-available model (\textsc{Llama-2}) on astrophysics-specific textual data from the arXiv.

In this paper we describe \text{PAPERCLIP} (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pre-training\footnote{Technically we fine tune rather than pre train, but ``PAPERCLIFT'' was rejected by the senior author of this paper.}), a method that connects astronomical image observations with natural language by leveraging the association between abstracts of successful observing proposals and images corresponding to downstream observations. 
%
Concretely, here we showcase the method using observations imaged by the \hubble Space Telescope (HST).
%
We show that fine-tuning a pre-trained CLIP ~\citep[Contrastive Language-Image Pre-training; ][]{radford2021learning} image-text model on observation-abstract pairs results in meaningful joint representations through quantitative and qualitative evaluation tests.
%
The method opens up the possibility of interacting with astronomical survey data using free-form natural language as an interface, which is a cornerstone of the success of the modern foundation model paradigm.
%

The CLIP family of foundation models, which in their original form embed images and associated captions into a common representation space via contrastive learning, have shown strong performance and generalization capabilities on a variety of downstream tasks including zero-shot classification and image retrieval.
%
The concept of associating diverse modalities via contrastive training has been employed in other scientific domains~\citep[e.g.,][]{liu2023text,Sanchez-Fernandez2022.11.17.516915,lanusse2023astroclip,cepeda2023geoclip}, and has been shown to be effective in learning semantically meaningful joint representations. Here we present for the first time an application associating astronomical data with the text modality.

@@

The paper is organized as follows.
%
In Sec.~\ref{sec:dataset}, we describe the \hubble dataset used in this work, including the curation and processing of observations as well as text captions.
%
In Sec.~\ref{sec:methodology}, we describe the methodology used to train and evaluate the model.
%
In Sec.~\ref{sec:results}, we present quantitative and qualitative results of our experiments on retrieval tasks.
%
We discuss future prospects and conclude in Sec.~\ref{sec:conclusion}.

@@

\begin{table}[h!]
  \centering
  \begin{tabular}{m{0.21\textwidth} p{1.9cm} p{1.9cm} m{8cm}}
      \toprule
      \centering \bfseries \hubble image & \centering \bfseries Obs. cycle \\ (Year) & \centering \bfseries Prop. ID & \centering \bfseries Proposal abstract (clipped) \tabularnewline
      \midrule
      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_2.pdf} & \centering \input{\thedatafolder/cycle_2.txt} \\ (1999) & \centering \input{\thedatafolder/id_2.txt} &  {\scriptsize \input{\thedatafolder/abs1_2.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_1.pdf} & \centering \input{\thedatafolder/cycle_1.txt} \\ (2013) & \centering \input{\thedatafolder/id_1.txt} &  {\scriptsize \input{\thedatafolder/abs1_1.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_3.pdf} & \centering \input{\thedatafolder/cycle_3.txt} \\ (2016) & \centering \input{\thedatafolder/id_3.txt} &  {\scriptsize \input{\thedatafolder/abs1_3.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_0.pdf} & \centering \input{\thedatafolder/cycle_0.txt} \\ (2019) & \centering \input{\thedatafolder/id_0.txt} &  {\scriptsize \input{\thedatafolder/abs1_0.txt}} \tabularnewline
      \bottomrule
  \end{tabular}
  \caption{Examples of \hubble images (left-most column) and corresponding clipped proposal abstracts (right-most column). The observation cycle and corresponding year, as well as proposal ID are shown in the second and third columns, respectively. The proposal IDs link to the Mikulski Archive for Space Telescopes (MAST) page corresponding to the proposal.}
  \label{tab:dataset}
\end{table}

@@

In total 31,859 images corresponding to 4,438 abstracts are included in the fine-tuning dataset.
%
3,194 images are held out for validation, with no abstract being common between training and validation sets in order to ensure an independent set of image-text pairs for evaluation. The held out images correspond to 438 unique abstracts.  % Double check numbers in final version.

@@

The model is evaluated by tracking the contrastive loss in Eq.~\eqrefb{eq:softmax_loss} as well as the top-$k\%$ retrieval accuracy on the held out validation set over the course of training.
%
The retrieval accuracy is defined as the fraction of associated captions which fall within the top $k\%$ of captions by cosine similarity of the normalized image and caption embeddings, averaged over the images in the validation set:
\begin{equation}
\text{Retrieval accuracy}_k = \frac{1}{|\mathcal V|} \sum_{i=1}^{|\mathcal V|} \mathbbm{1}\left[\operatorname{rank}\left({x}_i \cdot {y}_{i}; \{{x}_i \cdot {y}_{j}\}_{j=1}^{|\mathcal V|}\right) \leq \left\lfloor\frac{k}{100}|\mathcal V|\right\rfloor\right]
\label{eq:retrieval_accuracy}
\end{equation}
where $|\mathcal V|$ is the total number of images in the validation set, $\mathbbm{1}(\cdot)$ is the indicator function that returns 1 if the condition inside the brackets is true and 0 otherwise, $\operatorname{rank}\left({x}_i \cdot {y}_{i}; \{{x}_i \cdot {y}_{j}\}_{j=1}^{|\mathcal V|}\right)$ is a function that returns the rank of the cosine similarity between ${x}_i$ and ${y}_{i}$ among the cosine similarities between ${x}_i$ and all captions ${y}_j$ in the validation set, and $k$ is the percentage of top captions considered for the retrieval accuracy. Note that this metric is symmetric in the image and text modalities.

We also qualitatively evaluate the learned embeddings through image retrieval (i.e., retrieving the most relevant images from the validation set using natural language queries) and description retrieval (i.e., querying the astrophysical object classes and science use cases most relevant to a given observation, akin to zero-shot classification) experiments. 
%
For the description/text retrieval evaluation, we curate a list of possible text associations (i.e., classes) by querying the \textsc{Claude 2}\footnote{\url{https://claude.ai/}} large language model, which we show in App.~\ref{app:categories}.

@@

The model trained using LLM-summarized abstracts (red lines) is seen to perform slightly worse than the model using raw abstracts as captions (blue lines), despite the curation of the summarized-abstract dataset intended to provide a stronger image-text association signal.
%
Fine-tuning a small MLP head over frozen vision and text backbones (dotted green lines) and training from scratch (yellow lines) show a non-trivial improvement compared to the base model, although with deteriorated performance compared to fine-tuning with either summarized or raw abstracts.

\begin{figure*}[!h]
\includegraphics[width=0.99\textwidth]{plots/val_metrics.pdf}
\caption{The CLIP contrastive loss from Eq.~\eqrefb{eq:softmax_loss} (left) and the top-10\% retrieval accuracy from Eq.~\eqrefb{eq:retrieval_accuracy} (right) computed on the validation set over the course of training. Shown for the dataset with summarized abstracts as captions (red), dataset using raw proposal abstracts as captions (blue), only fine-tuning a small MLP head (dotted green), training from scratch (purple), and trained with shuffled image-text pairs (dashed orange).} 
\label{fig:retrieval_acc}
\end{figure*}

@@

Figure~\ref{fig:sim_valtrain} (right) shows the retrieval accuracy, as defined in Eq.~\eqrefb{eq:retrieval_accuracy}, as a function of the retrieval fraction $k$.
%
In this case, we evaluate all four models (fine tuned on raw abstracts (blue), fine-tuned on LLM-summarized abstracts (red), trained on LLM-summarized abstracts from scratch (yellow), and the base model (purple)) on the same captions dataset -- the summarized abstracts -- for a direct comparison.
%
Remarkably, the model trained on raw abstracts shows very similar performance when evaluated on the summarized abstracts compared to that trained on the summarized abstracts themselves, indicating that \emph{(1)} the image-text association signal is preserved in the summarization process, and \emph{(2)} the model is able to effectively leverage meaningful concepts in the noisy raw abstracts through weak supervision.

\begin{figure*}[!h]
  \includegraphics[width=0.49\textwidth]{plots/sim_val.pdf}
  % \includegraphics[width=0.45\textwidth]{plots/sim_summ1.pdf}
  \includegraphics[width=0.49\textwidth]{plots/retrieval.pdf}
  \caption{(Left) Distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$, shown when using the base CLIP model (purple lines), and the summary fine-tuned CLIP model (red line). Dashed lines correspond to models evaluated on image-text pairs with associations shuffled. (Right) Retrieval accuracy as a function of the retrieval fraction $k$ for the fine-tuned model on the summarized abstracts (red), fine-tuned on raw abstracts (blue), trained on LLM-summarized abstracts from scratch (yellow), and the base model (purple).}
  \label{fig:sim_valtrain}
  \end{figure*}

@@

The second row (an image of supernova 1987A) highlights an interesting pattern -- the base model erroneously attributes the object at the center of the image to a gravitational lens, while the fine-tuned model correctly identifies it as a supernova remnant. The kind of reasonable misattribution is common when querying the base model, and largely absent in the fine-tuned model.

@@

\begin{table}[h!]
  \centering
  \begin{tabular}{m{3cm} p{3cm} p{3cm} p{3cm} p{3cm}}
      \toprule
      \centering \bfseries Query & \multicolumn{4}{c}{\bfseries{Top-4 most similar images using \textcolor{deeppurple}{base CLIP model}}} \tabularnewline
      \midrule
      \texttt{\input{\thedatafolder/query_tti_base_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_0.pdf} \\ \input{\thedatafolder/propid_tti_base_1_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_1.pdf} \\ \input{\thedatafolder/propid_tti_base_1_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_2.pdf} \\ \input{\thedatafolder/propid_tti_base_1_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_3.pdf} \\ \input{\thedatafolder/propid_tti_base_1_3.txt}  \tabularnewline
      \midrule
       \texttt{\input{\thedatafolder/query_tti_base_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_0.pdf} \\ \input{\thedatafolder/propid_tti_base_0_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_1.pdf} \\ \input{\thedatafolder/propid_tti_base_0_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_2.pdf} \\ \input{\thedatafolder/propid_tti_base_0_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_3.pdf} \\ \input{\thedatafolder/propid_tti_base_0_3.txt}  \tabularnewline
      \midrule
      \texttt{\input{\thedatafolder/query_tti_base_2.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_0.pdf} \\ \input{\thedatafolder/propid_tti_base_2_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_1.pdf} \\ \input{\thedatafolder/propid_tti_base_2_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_2.pdf} \\ \input{\thedatafolder/propid_tti_base_2_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_3.pdf} \\ \input{\thedatafolder/propid_tti_base_2_3.txt}  \tabularnewline
      \midrule
      \texttt{\input{\thedatafolder/query_tti_base_3.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_0.pdf} \\ \input{\thedatafolder/propid_tti_base_3_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_1.pdf} \\ \input{\thedatafolder/propid_tti_base_3_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_2.pdf} \\ \input{\thedatafolder/propid_tti_base_3_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_3.pdf} \\ \input{\thedatafolder/propid_tti_base_3_3.txt}  \tabularnewline
      \bottomrule
  \end{tabular}
  \caption{For four text queries (left-most column), the four most similar images from the validation dataset by cosine similarity when using the \textbf{\textcolor{deeppurple}{base CLIP model}} (CLIP-ViT-B/16). The proposal ID associated with each image is given below the image and contains a hyperlink to the MAST page corresponding to the proposal.}
  \label{tab:tti_base}
\end{table}

@@

\begin{table}[h!]
  \centering
  \begin{tabular}{m{3cm} p{3cm} p{3cm} p{3cm} p{3cm}}
      \toprule
      \centering \bfseries Query & \multicolumn{4}{c}{\bfseries{Top-4 most similar images using \textcolor{deepred}{summary fine-tuned CLIP model}}} \tabularnewline
      \midrule
      \texttt{\input{\thedatafolder/query_tti_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_0.pdf} \\ \input{\thedatafolder/propid_tti_1_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_1.pdf} \\ \input{\thedatafolder/propid_tti_1_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_2.pdf} \\ \input{\thedatafolder/propid_tti_1_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_3.pdf} \\ \input{\thedatafolder/propid_tti_1_3.txt}  \tabularnewline
      \midrule
       \texttt{\input{\thedatafolder/query_tti_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_0_0.pdf} \\ \input{\thedatafolder/propid_tti_0_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_0_1.pdf} \\ \input{\thedatafolder/propid_tti_0_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_0_2.pdf} \\ \input{\thedatafolder/propid_tti_0_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_0_3.pdf} \\ \input{\thedatafolder/propid_tti_0_3.txt}  \tabularnewline
      \midrule
      \texttt{\input{\thedatafolder/query_tti_2.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_2_0.pdf} \\ \input{\thedatafolder/propid_tti_2_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_2_1.pdf} \\ \input{\thedatafolder/propid_tti_2_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_2_2.pdf} \\ \input{\thedatafolder/propid_tti_2_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_2_3.pdf} \\ \input{\thedatafolder/propid_tti_2_3.txt}  \tabularnewline
      \midrule
      \texttt{\input{\thedatafolder/query_tti_3.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_3_0.pdf} \\ \input{\thedatafolder/propid_tti_3_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_3_1.pdf} \\ \input{\thedatafolder/propid_tti_3_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_3_2.pdf} \\ \input{\thedatafolder/propid_tti_3_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_3_3.pdf} \\ \input{\thedatafolder/propid_tti_3_3.txt}  \tabularnewline
      \bottomrule
  \end{tabular}
  \caption{Same as Tab.~\ref{tab:tti_base}, but using the \textbf{\textcolor{deepred}{summary fine-tuned CLIP model}}.}
  \label{tab:tti}
\end{table}

>>>>

\def\preprintno{XXXX \SM{Requested from Charles}} % Insert correct preprint number

@@

There is considerable interest in developing custom foundation models for the sciences \citep[e.g., ][]{batatia2023foundation,subramanian2023towards}, with astrophysics being ripe for such an effort given the large amounts of publicly-available data and diverse modes of interacting with it.
%
The multi-modality inherent to astrophysical observations, with different types of data (e.g., images, spectra, light curves, textual descriptions) often available for a given target object, presents a unique opportunity.
%
This multi-modality was recently exploited in \textsc{AstroCLIP}~\citep{lanusse2023astroclip} to construct a joint embedding space between multi-band images and optical spectra from the Dark Energy Spectroscopic Instrument (DESI).
%
\textsc{AstroLLaMA}~\citep{nguyen2023astrollama,perkowski2024astrollama} is another recent effort to fine-tune a publicly-available model (\textsc{Llama-2}) on astrophysics-specific textual data from the arXiv.

In this paper, we describe \text{PAPERCLIP} (Proposal Abstracts Provide an Effective Representation \SM{I like `Representation' rather than `Refinement', I think the former emphasizes the central abstracts-as-captions aspect a bit more} for Contrastive Language-Image Pre-training\footnote{Technically, we fine tune rather than pre train, but ``PAPERCLIFT'' was rejected by the senior author of this paper.}), a method that connects astronomical image observations with natural language by leveraging the association between abstracts of successful observing proposals and images corresponding to downstream observations. 
%
Concretely, here we showcase the method using observations imaged by the \hubble Space Telescope (HST).
%
We show that fine-tuning a pre-trained CLIP ~\citep[Contrastive Language-Image Pre-training; ][]{radford2021learning} image-text model on observation-abstract pairs results in meaningful joint representations through quantitative and qualitative evaluation tests.
%
Our method opens up the possibility of interacting with astronomical survey data using free-form natural language as an interface, which is a cornerstone of the success of the modern foundation model paradigm.
%

The CLIP family of foundation models, which in their original form embed images and associated captions into a common representation space via contrastive learning, have shown strong performance and generalization capabilities on a variety of downstream tasks including zero-shot classification and image retrieval.
%
The concept of associating diverse modalities via contrastive training has been employed in other scientific domains~\citep[e.g.,][]{liu2023text,Sanchez-Fernandez2022.11.17.516915,lanusse2023astroclip,cepeda2023geoclip}, and has been shown to be effective in learning semantically meaningful joint representations. Here, we present for the first time an application associating astronomical data with the text modality.

@@

The rest of this paper is organized as follows.
%
In Sec.~\ref{sec:dataset}, we describe the \hubble dataset used in this work, including the curation and processing of observations as well as text captions.
%
In Sec.~\ref{sec:methodology}, we describe the methodology used to train and evaluate the model.
%
In Sec.~\ref{sec:results}, we present quantitative and qualitative results of our experiments on retrieval tasks.
%
We discuss future prospects and conclude in Sec.~\ref{sec:conclusion}.

@@

\begin{table}[h!]
  \centering
  \begin{tabular}{m{0.20\textwidth} p{1.9cm} p{1.9cm} m{7.5cm}}
      \toprule
      \centering \bfseries \hubble image & \centering \bfseries Obs. cycle \\ (Year) & \centering \bfseries Prop. ID & \centering \bfseries Proposal abstract (clipped) \tabularnewline
      \midrule
      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_2.pdf} & \centering \input{\thedatafolder/cycle_2.txt} \\ (1999) & \centering \input{\thedatafolder/id_2.txt} &  {\scriptsize \input{\thedatafolder/abs1_2.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_1.pdf} & \centering \input{\thedatafolder/cycle_1.txt} \\ (2013) & \centering \input{\thedatafolder/id_1.txt} &  {\scriptsize \input{\thedatafolder/abs1_1.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_3.pdf} & \centering \input{\thedatafolder/cycle_3.txt} \\ (2016) & \centering \input{\thedatafolder/id_3.txt} &  {\scriptsize \input{\thedatafolder/abs1_3.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_0.pdf} & \centering \input{\thedatafolder/cycle_0.txt} \\ (2019) & \centering \input{\thedatafolder/id_0.txt} &  {\scriptsize \input{\thedatafolder/abs1_0.txt}} \tabularnewline
      \bottomrule
  \end{tabular}
  \caption{Examples of \hubble images (left-most column) and corresponding clipped proposal abstracts (right-most column). The observation cycle and corresponding year, as well as proposal ID, are shown in the second and third columns, respectively. The proposal IDs link to the Mikulski Archive for Space Telescopes (MAST) page corresponding to the proposal.}
  \label{tab:dataset}
\end{table}

@@

In total, 31,859 images corresponding to 4,438 abstracts are included in the fine-tuning dataset.
%
3,194 images are held out for validation, with no abstract being common between training and validation sets in order to ensure an independent set of image-text pairs for evaluation. The held out images correspond to 438 unique abstracts.  % Double check numbers in final version.

@@

The model is evaluated by tracking the contrastive loss in Eq.~\eqrefb{eq:softmax_loss} as well as the top-$k\%$ retrieval accuracy on the held out validation set over the course of training.
%
The retrieval accuracy is defined as the fraction of associated captions which fall within the top $k\%$ of captions by cosine similarity of the normalized image and caption embeddings, averaged over the images in the validation set:
\begin{equation}
\text{Retrieval accuracy}_k = \frac{1}{|\mathcal V|} \sum_{i=1}^{|\mathcal V|} \mathbbm{1}\left[\operatorname{rank}\left({x}_i \cdot {y}_{i}; \{{x}_i \cdot {y}_{j}\}_{j=1}^{|\mathcal V|}\right) \leq \left\lfloor\frac{k}{100}|\mathcal V|\right\rfloor\right]
\label{eq:retrieval_accuracy}
\end{equation}
where $|\mathcal V|$ is the total number of images in the validation set, $\mathbbm{1}[\cdot]$ is the indicator function that returns 1 if the condition inside the brackets is true and 0 otherwise, $\operatorname{rank}\left({x}_i \cdot {y}_{i}; \{{x}_i \cdot {y}_{j}\}_{j=1}^{|\mathcal V|}\right)$ is a function that returns the rank of the cosine similarity between ${x}_i$ and ${y}_{i}$ among the cosine similarities between ${x}_i$ and all captions ${y}_j$ in the validation set, and $k$ is the percentage of top captions considered for the retrieval accuracy. Note that this metric is symmetric in the image and text modalities.

We also qualitatively evaluate the learned embeddings through image retrieval (i.e., retrieving the most relevant images from the validation set using natural language queries) and description retrieval (i.e., querying the astrophysical object classes and science use cases most relevant to a given observation, akin to zero-shot classification) experiments. 
%
For the description/text retrieval evaluation, we define a list of possible text associations (i.e., classes), which we show in App.~\ref{app:categories}, by querying the \textsc{Claude 2}\footnote{\url{https://claude.ai/}} large language model along with manual curation.

@@

The model trained using LLM-summarized abstracts (red lines) is seen to perform slightly worse than the model using raw abstracts as captions (blue lines), despite the curation of the summarized-abstract dataset intended to provide a stronger image-text association signal.
%
Fine-tuning a small MLP head over frozen vision and text backbones (dotted green lines) and training from scratch with summarized abstracts as captions (yellow lines) show a non-trivial improvement compared to the base model, although with deteriorated performance compared to fine-tuning with either summarized or raw abstracts.

\begin{figure*}[!h]
\includegraphics[width=0.99\textwidth]{plots/val_metrics.pdf}
\caption{The CLIP contrastive loss from Eq.~\eqrefb{eq:softmax_loss} (left) and the top-10\% retrieval accuracy from Eq.~\eqrefb{eq:retrieval_accuracy} (right) computed on the validation set over the course of training. Shown for the dataset with summarized abstracts as captions (red), dataset using raw proposal abstracts as captions (blue), only fine-tuning a small MLP head (dotted green), training from scratch with summarized abstracts as captions (yellow), and trained with shuffled image-text pairs (dashed orange).} 
\label{fig:retrieval_acc}
\end{figure*}

@@

Figure~\ref{fig:sim_valtrain} (right) shows the retrieval accuracy, as defined in Eq.~\eqrefb{eq:retrieval_accuracy}, as a function of the retrieval fraction $k\%$.
%
In this case, we evaluate all four models (fine tuned on raw abstracts (blue), fine-tuned on LLM-summarized abstracts (red), trained on LLM-summarized abstracts from scratch (yellow), and the base model (purple)) on the same captions dataset -- the summarized abstracts -- for a direct comparison.
%
Remarkably, the model trained on raw abstracts shows very similar performance when evaluated on the summarized abstracts compared to that trained on the summarized abstracts themselves, indicating that \emph{(1)} the image-text association signal is preserved in the summarization process, and \emph{(2)} the model is able to effectively leverage meaningful concepts in the noisy raw abstracts through weak supervision.

\begin{figure*}[!h]
  \includegraphics[width=0.49\textwidth]{plots/sim_val.pdf}
  % \includegraphics[width=0.45\textwidth]{plots/sim_summ1.pdf}
  \includegraphics[width=0.49\textwidth]{plots/retrieval.pdf}
  \caption{(Left) Distribution of cosine similarities between corresponding image and text embeddings, $x_i$ and $y_i$, shown when using the base CLIP model (purple lines), and the summary fine-tuned CLIP model (red line). Dashed lines correspond to models evaluated on image-text pairs with associations shuffled. (Right) Retrieval accuracy as a function of the retrieval fraction $k$ for the fine-tuned model on the summarized abstracts (red), fine-tuned on raw abstracts (blue), trained on summarized abstracts from scratch (yellow), and the base model (purple).}
  \label{fig:sim_valtrain}
  \end{figure*}

@@

The second row (an image of supernova 1987A) highlights an interesting pattern -- the base model erroneously attributes the object at the center of the image to a gravitational lens, while the fine-tuned model correctly identifies it as a supernova remnant. This kind of reasonable misattribution is common when querying the base model, and largely absent in the fine-tuned model.

@@

\begin{table}[h!]
  \centering
  \begin{tabular}{m{2.7cm} p{2.9cm} p{2.9cm} p{2.9cm} p{2.9cm}}
      \toprule
      \centering \bfseries Query & \multicolumn{4}{c}{\bfseries{Top-4 most similar images using \textcolor{deeppurple}{base CLIP model}}} \tabularnewline
      \midrule
      \texttt{\input{\thedatafolder/query_tti_base_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_0.pdf} \\ \input{\thedatafolder/propid_tti_base_1_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_1.pdf} \\ \input{\thedatafolder/propid_tti_base_1_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_2.pdf} \\ \input{\thedatafolder/propid_tti_base_1_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_3.pdf} \\ \input{\thedatafolder/propid_tti_base_1_3.txt}  \tabularnewline
      \midrule
       \texttt{\input{\thedatafolder/query_tti_base_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_0.pdf} \\ \input{\thedatafolder/propid_tti_base_0_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_1.pdf} \\ \input{\thedatafolder/propid_tti_base_0_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_2.pdf} \\ \input{\thedatafolder/propid_tti_base_0_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_3.pdf} \\ \input{\thedatafolder/propid_tti_base_0_3.txt}  \tabularnewline
      \midrule
      \texttt{\input{\thedatafolder/query_tti_base_2.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_0.pdf} \\ \input{\thedatafolder/propid_tti_base_2_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_1.pdf} \\ \input{\thedatafolder/propid_tti_base_2_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_2.pdf} \\ \input{\thedatafolder/propid_tti_base_2_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_3.pdf} \\ \input{\thedatafolder/propid_tti_base_2_3.txt}  \tabularnewline
      \midrule
      \texttt{\input{\thedatafolder/query_tti_base_3.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_0.pdf} \\ \input{\thedatafolder/propid_tti_base_3_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_1.pdf} \\ \input{\thedatafolder/propid_tti_base_3_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_2.pdf} \\ \input{\thedatafolder/propid_tti_base_3_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_3.pdf} \\ \input{\thedatafolder/propid_tti_base_3_3.txt}  \tabularnewline
      \bottomrule
  \end{tabular}
  \caption{For four text queries (left-most column), the four most similar images from the validation dataset by cosine similarity when using the \textbf{\textcolor{deeppurple}{base CLIP model}} (CLIP-ViT-B/16). The proposal ID associated with each image is given below the image and contains a hyperlink to the MAST page corresponding to the proposal.}
  \label{tab:tti_base}
\end{table}

@@

\begin{table}[h!]
  \centering
  \begin{tabular}{m{2.7cm} p{2.9cm} p{2.9cm} p{2.9cm} p{2.9cm}}
      \toprule
      \centering \bfseries Query & \multicolumn{4}{c}{\bfseries{Top-4 most similar images using \textcolor{deepred}{summary fine-tuned CLIP model}}} \tabularnewline
      \midrule
      \texttt{\input{\thedatafolder/query_tti_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_0.pdf} \\ \input{\thedatafolder/propid_tti_1_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_1.pdf} \\ \input{\thedatafolder/propid_tti_1_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_2.pdf} \\ \input{\thedatafolder/propid_tti_1_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_1_3.pdf} \\ \input{\thedatafolder/propid_tti_1_3.txt}  \tabularnewline
      \midrule
       \texttt{\input{\thedatafolder/query_tti_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_0_0.pdf} \\ \input{\thedatafolder/propid_tti_0_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_0_1.pdf} \\ \input{\thedatafolder/propid_tti_0_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_0_2.pdf} \\ \input{\thedatafolder/propid_tti_0_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_0_3.pdf} \\ \input{\thedatafolder/propid_tti_0_3.txt}  \tabularnewline
      \midrule
      \texttt{\input{\thedatafolder/query_tti_2.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_2_0.pdf} \\ \input{\thedatafolder/propid_tti_2_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_2_1.pdf} \\ \input{\thedatafolder/propid_tti_2_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_2_2.pdf} \\ \input{\thedatafolder/propid_tti_2_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_2_3.pdf} \\ \input{\thedatafolder/propid_tti_2_3.txt}  \tabularnewline
      \midrule
      \texttt{\input{\thedatafolder/query_tti_3.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_3_0.pdf} \\ \input{\thedatafolder/propid_tti_3_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_3_1.pdf} \\ \input{\thedatafolder/propid_tti_3_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_3_2.pdf} \\ \input{\thedatafolder/propid_tti_3_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_3_3.pdf} \\ \input{\thedatafolder/propid_tti_3_3.txt}  \tabularnewline
      \bottomrule
  \end{tabular}
  \caption{Same as Tab.~\ref{tab:tti_base}, but using the \textbf{\textcolor{deepred}{summary fine-tuned CLIP model}}.}
  \label{tab:tti}
\end{table}
